---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
d <- read.csv("activity.csv",header=TRUE,colClasses=c("integer","Date","integer"))
d["i_factor"] = as.factor(d$interval)
d$wd = weekdays(d$date)
d$weekend = "weekday"
d$weekend[(d$wd == "Sonntag") | (d$wd=="Samstag")] = "weekend"
d$weekend = as.factor(d$weekend)
d_agg_sum <- aggregate(d[1],list(d$date),FUN=sum, na.rm=TRUE)
```

## What is mean total number of steps taken per day?
You'll find the total number of steps taken per day in the following table.

```{r, results='asis'}
library(xtable)
xt <- xtable(d_agg_sum)
print(xt, type="html")
```

A histogram of the total number of steps taken each day:
```{r}
hist(d_agg_sum$steps)
```

The mean of the total number of steps taken per day is `r mean(d_agg_sum$steps)`.

The median of the total number of steps taken per day is `r median(d_agg_sum$steps)`.

## What is the average daily activity pattern?

```{r}
ts <- aggregate(d$steps,list(d$interval),FUN=mean,na.rm=TRUE)
plot(ts,type='l',xlab = "Interval",ylab = "avg steps across all days")
```

Interval `r ts[ts[2]==max(ts[2]),][1]` contains the maximum average.



## Imputing missing values

There are `r nrow(d)-nrow(na.omit(d))` rows containing missing values.

The missing values are filled with the mean value of that interval across all days in the dataset.

```{r}
for (i in 1:nrow(d)) {
  if (is.na(d$steps[i])) {
    # interval id
    iv <- d$interval[i]
    
    # mean for this interval
    steps_mean <- ts[ts[1]==iv]
    
    d$steps[i]<-steps_mean[2]
  }
}
```

A histogram of the total number of steps taken each day:

```{r}
#date_steps_imp <- aggregate(steps ~ date, d, sum)
date_steps_imp <- aggregate(d[1],list(d$date),FUN=sum)
hist(date_steps_imp$steps)


```

Mean:
```{r}
# Mean of steps per day imputed
mean(date_steps_imp$steps)

# Mean of steps per day (w/ NA)
mean(d_agg_sum$steps)
```
The mean is raised by ~1500 steps

Median:
```{r}
# Median of steps per day imputed
median(date_steps_imp$steps)

# Median of steps per day (w/ NA)
median(d_agg_sum$steps)
```
Median is raised.


## Are there differences in activity patterns between weekdays and weekends?

```{r}
ts_imputed <- aggregate(steps ~ interval+weekend, d, mean)

# make the panel plot for weekdays and weekends
library(ggplot2)

qplot(interval, steps, data=ts_imputed, geom=c("line"), xlab="Interval", 
      ylab="Number of steps", main="") + facet_wrap(~ weekend, ncol=1)
```
